---
title: "Challenger O-Ring Data -- Logistic Regression"
output: 
  html_document: 
    theme: cerulean
    code_folding: hide
---

<hr />

#### Background

<a href="http://www.history.com/topics/challenger-disaster/videos/engineering-disasters---challenger"> Watch the Introductory Movie </a>

```{r, include=FALSE}
library(alr3)
library(mosaic)
library(DT)
```

```{r}
datatable(challeng)
```

As described in the video, it was known that the o-rings could fail when the tempurature was below 53&deg; F. It then stated that "Nasa had already launched shuttles below 53&deg; and gotten away with it." This second statement is actually false. The lowest temperature of any of the 23 prior launches (before the Challenger explosion) was 53&deg; F [[view source]](http://www.ics.uci.edu/~staceyah/111-202/handouts/Dalal_etal_1989-Challenger.pdf). The "evidence" that the o-rings could fail below 53&deg; was based on a simple conclusion that since the launch at 53&deg; experienced two o-ring failures, it seemed unwise to launch below that temperature. The statement that NASA had launched previously with o-ring failures occurring, but the launch still being successful was true, but all launches had a temperature of at least 53&deg; F. Thus, whether or not temperature was the root cause of the failures was debatable. 

<br />

#### Logistic Model

To model the probability of the o-rings failing at various temperatures, we could apply the logistic regression model
$$
  P(Y_i = 1|x_i) = \frac{e^{\beta_0+\beta_1 x_i}}{1+e^{\beta_0 + \beta_1 x_i}} = \pi_i
$$
where for observation $i$: 

* $Y_i = 1$ denotes at least one o-ring failing for a given launch, 
* $Y_i=0$ denotes no o-rings failing (successful launch), and 
* $x_i$ denotes the outside temperature in degrees farenheit at the time of the launch. 

Note that if $\beta_1$ is zero in the above model, then $x_i$ (temperature) provides no insight about the probability of a failed O-ring. Thus, we could test the hypothesis that

$$
  H_0: \beta_1 = 0 \\
  H_a: \beta_1 \neq 0
$$

<br />

#### Fitting the Model

To obtain estimates of the coefficients $\beta_0$ and $\beta_1$ for the `challeng` data, we apply the following R code.

```{r}
chall.glm <- glm(Fail>0 ~ Temp, data=challeng, family=binomial)
summary(chall.glm)
```

Thus the estimated model for $\pi_i$ is given by
$$
  P(Y_i = 1|x_i) \approx \frac{e^{15.043-0.232 x_i}}{1+e^{15.043 - 0.232 x_i}} = \hat{\pi}_i
$$
where $b_0 = 15.043$ is the value of the `(Intercept)` which estimates $\beta_0$ and $b_1 = -0.232$ is the value of `Temp` which estimates $\beta_1$.

Importantly, the $p$-value for the test of `Temp` shows a significant result $(p = 0.0320)$ giving sufficient evidence to conclude that $\beta_1 \neq 0$, which allows us to conclude that temperature effects the probability of an O-ring failure.

<br />

#### Visualizing the Model

To visualize this simple logistic regression we could make the following plot.

```{r}
plot( Fail>0 ~ Temp, data=challeng, main="", ylab='Probability of O-rings Failing', pch=16)
curve(exp(15.043-0.232*x)/(1+exp(15.043-0.232*x)), from=53, to=85, add=TRUE)
```

<br />

#### Diagnosing the Model

To demonstrate that the logistic regression is a good fit to these data we apply the Hosmer-Lemeshow goodness of fit test (since there are only a couple repeated $x$-values) from the `library(ResourceSelection)`. 

```{r, warning=FALSE, message=FALSE}
library(ResourceSelection)
```
```{r}
hoslem.test(chall.glm$y, chall.glm$fitted, g=6)
# Note: doesn't give a p-value for g >= 7, default is g=10.
# Larger g is usually better than smaller g.
```
Recall that the null hypothesis is that the logistic regression is a good fit for the data, thus we conclude the null, and claim that the logistic regression is appropriate.

<br />

#### Interpreting the Model

Since the temperature being zero is not really realistic for this model, the value of $e^{b_0}$ is not interpretable. However, the value of $e^{b_1} = e^{-0.232} \approx 0.79$ shows that the odds of the o-rings failing for a given launch decreases by a factor of 0.79 for every 1&deg; F increase in temperature. (Note that this implies that every 1&deg; F decrease in temperature increases the odds of a failed o-ring by a factor of $e^{0.232} \approx 1.26$.) The Challenger shuttle was launched at a tempurature of 31&deg; F. By waiting until 53&deg; F, the odds of failure would have been decreased by a factor of $e^{-0.232(53-31)}\approx 0.006$. Note that for a temperature of 31&deg; F our model puts the probability of a failure at
$$
  P(Y_i = 1|x_i) \approx \frac{e^{15.043-0.232\cdot 31}}{1+e^{15.043 - 0.232 \cdot 31}} = \hat{\pi}_i
$$
which, using R to do this calculation we get $\hat{\pi_i} \approx$
```{r}
predict(chall.glm, data.frame(Temp=31), type='response')
```