---
title: "Rent"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---

**Week 1 Analysis**

```{r, include=FALSE}
library(mosaic)
library(pander)
library(tidyverse)
library(DT) 
Rent <- read_csv("../Data/Rent.csv")
```

<br />

## Background

Here is a data table showing the available approved housing apartment options at BYU-Idaho for single students. There are 122 entries comprising [insert how many] female and [insert how many] male apartment options.

```{r, warning=FALSE}
# Code to get you started. 
# Only the first 100 observations are shown here because the dataset is really too big to show all of it in a markdown file.
datatable(Rent, options=list(lengthMenu = c(3,10,30)), extensions="Responsive")
```


<!-- Write a few statements here that introduce the data table showing apartments for only your gender. -->

...delete this and type here...

[link name](www.linkaddress.com)

[Analysis Rubric](./Analyses/AnalysisRubric.html)
[Graphical Summaries](GraphicalSummaries.html)
[Stephanie](./Analyses/Stephanie.html)
[Boxplots](GraphicalSummaries.html#boxplots)


```{r}
# Use this R-chunk to...
# ...filter this data set down to show only your gender's apartment options. Use the R-chunk below to do this. Call your new data set `Rentf` if you are a female or `Rentm` if you are a male. Then show the new data table using the `datatable(...)` function. -->


```


<!-- Write a few statements here that introduce your graphic and what it shows. -->

...delete this and type here...


```{r}
# Use this R-chunk to...
# ...create a graphic that shows how your apartment (or a friend's) compares to all of the options available.


```

<!-- Write a few statements here that introduce your table of numerical summaries and what they show. -->

...delete this and type here...


```{r}
# Use this R-chunk to...
# ...compute and display a meaningful table of numerical summaries supporting your above graphic.


```


## Conclusion

...delete this and type an interesting final conclusion here...



