---
title: "Normal Probability Plot"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, include=FALSE}
library(mosaic)
library(car)
```

----

There are four main trends that occur in a normal probability plot. Examples of each are plotted below with a histogram of the data next to the normal probability plot.

Often the plot is called a Q-Q Plot, which stands for quantile-quantile plot. The idea is to compare the observed distribution of data to what the distribution should look like in theory if it was normal. Q-Q Plots are more general than normal probability plots because they can be used with any theoretical distribution, not just the normal distribution.  

```{r}

par(mfrow=c(2,2),  mai=c(.5,.5,.25,.25), mgp=c(1,.75,0))

set.seed(123)

  tmp <- rnorm(100)
  qqnorm(tmp, pch=20, ylab="Observed", xaxt='n', yaxt='n')
  qqline(tmp)
  hist(tmp, xlab="", xaxt='n', yaxt='n', main="Normal")
  
  tmp <- Davis$weight
  qqnorm(tmp, pch=20, ylab="Observed", xaxt='n', yaxt='n')
  qqline(tmp)
  hist(tmp, xlab="", xaxt='n', yaxt='n', main="Right-skewed",
       breaks=15)
  
par(mfrow=c(2,2),  mai=c(.5,.5,.25,.25), mgp=c(1,.75,0))

  tmp <- rbeta(100, 5,1)
  qqnorm(tmp, pch=20, ylab="Observed", xaxt='n', yaxt='n')
  qqline(tmp)
  hist(tmp, xlab="", xaxt='n', yaxt='n', main="Left-skewed",
       breaks=seq(min(tmp),max(tmp), length.out=13))
  
  tmp <- rbeta(100,2,2)
  qqnorm(tmp, pch=20, ylab="Observed", xaxt='n', yaxt='n')
  qqline(tmp)
  hist(tmp, xlab="", xaxt='n', yaxt='n', main="Heavy-tailed")

  
  
```
